---
title: "identifying duplicates and their position"
author: "Gakinya"
date: "August 25, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r}
#ls()

``{r}
ls()
rm(list = ls())
require(Hmisc)
dataIN16_2<-spss.get("C:\\Users\\user\\Google Drive\\TOURISMdatafinal_txt\\datain2016_6.sav")
attach(dataIN16_2)
detach(dataIN16_2)
names(dataIN16_2)
```
```

```{r}
attach(dataOUT2016)













names(dataOUT2016)
length(dataOUT2016$SERIAL_NUMBER)
b<-dataOUT2016$SERIAL_NUMBER
b
b1<-duplicated(b)
head(b1)
b2<-as.numeric(b1)
head(b2)
b3<-which(b2 %in% c(1))
b3

```


```{r}
names(dataOUT2016)
```

"DATE_OF_INTERVIEW"_DATE_OF_DEPARTURE"
```{r}
is.na(dataOUT2016$DATE_OF_INTERVIEW)## checking not applicable
is.na(dataOUT2016$Q4_DATE_OF_DEPARTURE)
```
```{r}
not_applicable_NA<-is.na(dataOUT2016$DATE_OF_INTERVIEW)
head(not_applicable_NA)
not_applicable_NA_numeric_covert<-as.numeric(not_applicable_NA)
head(not_applicable_NA_numeric_covert)
pos_NA_row1<-which(not_applicable_NA_numeric_covert %in% c(1))
pos_NA_row1
```

```{r}
options(max.print=5000)
```
```{r}
not_applicable_NA1<-is.na(dataOUT2016$Q4_DATE_OF_DEPARTURE)
head(not_applicable_NA1)
not_applicable_NA_numeric_covert1<-as.numeric(not_applicable_NA1)
head(not_applicable_NA_numeric_covert1)
pos_NA_row2<-which(not_applicable_NA_numeric_covert1 %in% c(1))
pos_NA_row2
```
```{r}
arr_char<-as.character(dataOUT2016$DATE_OF_INTERVIEW)
head(arr_char)
```
```{r}
dep_char<-as.character(dataOUT2016$Q4_DATE_OF_DEPARTURE)
head(dep_char)
```
```{r}
arr_date<-as.Date(arr_char,"%d%m%Y")
head(arr_date)
```
```{r}
dep_date<-as.Date(dep_char,"%d%m%Y")
head(dep_date)
```
```{r}
date_diff<- arr_date-dep_date
(date_diff)
```
```{r}
date_dates<-data.frame(dep_date,date_diff)
with(date_dates, count(date_diff))
```
```{r}
require(dplyr)
```

