---
title: "Data analysis,Hypothesis testing, Regression"
author: "Evelyn Gakinya"
date: "7/25/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,warning=FALSE,
                      fig.width = 8, fig.height = 6)
```


## The Document contains 6 sections: Section 4 contains data visualization,Section 5: Test Statistics, section 6: Linear regression
### 0.1 Loading the packages required
```{r}
###  create a vector of packages to be installed
library(tidyverse)
library(DT)
library(lubridate)
library(patchwork)
```
### 1. Read in the dataset

```{r}
library(readr)
df<-read.csv("C:\\Users\\EVAH\\Desktop\\RSessions\\Zindi\\Train_v2.csv",header = TRUE,sep=",")
```
### 2. Convert character variables to factor variables 
```{r}
## 2.1. bank_account
df <- df %>% 
  mutate(bank_account = fct_relevel(bank_account, "No", "Yes"))

## 2.2. location_type
df <- df %>% 
  mutate(location_type = fct_relevel(location_type, "Rural" , "Urban"))

## 2.3. cellphone_access
df <- df %>% 
  mutate(cellphone_access = fct_relevel(cellphone_access,  "No", "Yes"))

## 2.4. gender_of_respondent
df <- df %>% 
  mutate(gender_of_respondent = fct_relevel(gender_of_respondent,  "Female" ,"Male"))


## 2.5. relationship_with_head
df <- df %>% 
  mutate(relationship_with_head = fct_relevel(relationship_with_head,  
        "Child","Spouse", "Parent", "Head of Household",  
        "Other relative",   "Other non-relatives" ))

## 2.6. marital_status
df <- df %>% 
  mutate(marital_status = fct_relevel(marital_status,  
        "Single/Never Married","Divorced/Seperated","Widowed",
        "Married/Living together","Dont know"))

## 2.7. education_level
df <- df %>% 
  mutate(education_level = fct_relevel(education_level,  
        "No formal education", "Primary education", "Secondary education", 
        "Tertiary education", "Vocational/Specialised training","Other/Dont know/RTA"))

```
### 3. Generate 2 new variables i.e monthly income and date they registered for xyz
```{r}
 ## 3.1 monthly income

set.seed(2020)
income_values <- sample(c(5000 : 150000), nrow(df))
df <- df %>% 
  mutate(income = income_values,
         income = ifelse(job_type == "No Income", NA, income))
## 3.2 date of registering for xyz
date_vec <- sample(seq(as.Date('2014/01/01'), as.Date('2014/12/31'), by="day"), nrow(df), replace = T)
df <- df %>% 
  mutate(date = date_vec) 
## Generate the month
df <- df %>% 
  mutate(month = month(date, abbr = T, label = T)) 
```

### 4. Main graphs

#### 4.1 Bar graph

##### 4.1.1 Single categorical variable
Distribution of cellphone access
```{r}
## First generate a table
tab1 <-df %>% 
  group_by(cellphone_access) %>% 
  count() %>%  
  ungroup() %>% 
  mutate(Perc= round(n/sum(n) *100,0))

## Plot a graph
plot1 <- ggplot(data = tab1, aes(x = cellphone_access, y = Perc))+
  geom_bar(stat = "identity", fill ="maroon", width = 0.7)+
  geom_text(aes(label = Perc), size = 5, hjust = 0.5, vjust = -0.25)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, colour = "red"),
        axis.title = element_text(face = "italic", size = 10, colour = "black"),
        axis.text = element_text(face = "plain", size = 10, colour = "red"),
        axis.line = element_line(size = 1.5, colour = "green"),
        plot.subtitle = element_text(face = "italic", hjust = 0.5, colour = "brown"),
        plot.caption = element_text(face = "italic", size = 10, colour = "purple"))+
  labs(title = "Distribution of cellphone access",
       x = "",y = "Percentage",
       caption = " Prepared by Evelyn Gakinya")+
  ylim(c(0,100))
plot1
```
##### 4.1.2 Two categorical variables (main variable and grouping variable)

Distribution of bank account availability by country


```{r}
tab2 <- df %>% 
    group_by(location_type, bank_account) %>% 
    summarise(count2 = n()) %>% 
    mutate(percentage = round((count2/ sum(count2))*100 , 1)) 

plot2 <- ggplot(data=tab2, aes(x = location_type, y = percentage, fill = bank_account))+
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = percentage ), hjust=0.5, vjust=-0.3, size=3, position = position_dodge(width = 0.9))+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, colour = "red"),
        axis.title = element_text(face = "italic", size = 10, colour = "black"),
        axis.text = element_text(face = "plain", size = 10, colour = "red"),
        axis.line = element_line(size = 1.5, colour = "green"),
        plot.subtitle = element_text(face = "italic", hjust = 0.5, colour = "brown"),
        plot.caption = element_text(face = "italic", size = 10, colour = "purple"))+
  scale_fill_manual(values = c("black","yellow"))+
  labs(title =" Distribution of bank avalibility by location type", x = "location type",
       y = "percentage", subtitle = "2 variables",
       caption = "prepared by Evelyn Gakinya")
plot2
```
#### 4.2 Line graph

##### 4.2.1 Single categorical variable

Distribution of registrations by month
```{r}
tab5 <- df %>% 
  group_by(month) %>% 
  summarise(avg_age = round(mean(age_of_respondent, na.rm = T),1))

plot5 <- ggplot(data = tab5, aes(x = month, y = avg_age, group = 1))+
          geom_point()+
          geom_line(stat = "identity", size = 1, color = "maroon", linetype = "solid")+
          geom_text(aes(label = avg_age), vjust = 0, hjust = 1)+
         theme(plot.title = element_text(face = "bold", hjust = 0.5, colour = "red"),
        axis.title = element_text(face = "italic", size = 10, colour = "black"),
        axis.text = element_text(face = "plain", size = 10, colour = "red"),
        axis.line = element_line(size = 1.5, colour = "green"),
        plot.subtitle = element_text(face = "italic", hjust = 0.5, colour = "brown"),
        plot.caption = element_text(face = "italic", size = 10, colour = "purple"))+
        labs(title = "Average age of respondents registering per month",
             x="month", y ="Average Age",
             caption = "prepared by Evelyn Gakinya")+
  ylim(c(37.5,39.5))
plot5
```
##### 4.2.2 Two categorical variables (main variable and grouping variable)
```{r}
my_theme <-theme(axis.line = element_line(colour = "black", size = 1),
        panel.grid = element_blank(),
        panel.background = element_rect(fill = NA),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(hjust = 0.5, size = 18 ,family = "Source Sans Pro Semibold"),
        plot.caption = element_text(size = 14))

```

Distribution of registrations per month and country
```{r}
## table
tab6 <- df %>% 
  group_by(month, country) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(country) %>% 
  mutate(perc = round(n/sum(n)*100,1))

## graph
plot6 <- ggplot(data = tab6, aes(x = month, y = perc, group = country, color = country))+
          geom_point()+
          geom_line(stat = "identity", size = 1)+
          geom_text(aes(label = perc), hjust = 1.5, vjust = 0)+
          my_theme+
          scale_color_brewer(palette = "Spectral")
          labs(title = "Registartions by month",
                x = "month", y = "Percentage",
                caption = "prepared by Evelyn Gakinya")
              
plot6
```
### 5. Hypothesis testing

#### 5.1 Chi-Square Test
```{r}
##Example 2
##A child psychologist believes that children perform better on tests when they are given perceived
##freedom of choice. To test this belief, the psychologist carried out an experiment in which  200 
##third graders were randomly assigned to two groups,  A  and  B . Each child was given the same 
##simple logic test. However in group  B , each child was given the freedom to choose a text 
##booklet from many with various drawings on the covers. The performance of each child was rated as
##Very Good, Good, and Fair. The results are summarized in the table provided. Test, at the  5%  
##level of significance, whether there is sufficient evidence in the data to support the 
##psychologists belief.
##Step 1
 #### H0: Freedom of Choice and Perfomance  are independent
 
 #### H1:Freedom of Choice and Perfomance  are not independent
tab <- cbind(c(32, 55, 10), c(29, 61, 13))
dimnames(tab) <- list(Perfomance = c("Very Good", "Good", "Fair"),Group = c("A", 	"B"))
tab
```
##### 5.1.1 Determine the chi square test statistics
```{r}
chisq.test(tab, correct = F)
```
##### 5.1.2 Determine the chi- square tabulated

```{r}
qchisq(1-0.05,2)
```
##### 5.1.3 Make Conclusion
```{r}
##the calculated chi-square statistics = 0.669793 is less than tabulated = 5.991 thus falls in the
##non rejection region thus we fail to reject H0 and conclude that Freedom of Choice and 
##Perfomance  are independent
```


#### 5.2 t-test

```{r}
##Your company wants to improve sales. Past sales data indicate that the average sales was $100 per 
##transaction. After training your sales force, recent sales data (taken from a sample of 25 salesmen), 
##indicates an average sales of $130, with a standard deviation of $15. Did the training work? 
##Test your hypothesis at 5% alpha level. 
###Solution:
#one tailed test(area to the right)
improve_sales<- c(rnorm(25, mean = 130, sd = 15))
improve_sales

```

##### 5.2.1 Determine the t-statistics
```{r}
t.test(improve_sales, mu = 100, alternative = "greater")
```
##### 5.2.2 Make Conclusion
```{r}
####The p-value of the test is 1.081^{-10}, which is less than the significance level alpha = 0.05. 
####Thus we reject the null hypothesis and conclude that the average sale is greater than 100 thus thetraining worked.
```


### 6. Regression

#### 6.1 Linear Regression
##### 6.1.1  Reading in the data
```{r}
data2 <- read.csv("C:\\Users\\EVAH\\Desktop\\RSessions\\wafanyikazi.csv")
```
##### 6.1.2  Reading in the data
```{r}
## Fit a model that predicts income based on Gender, Role and Age
Linearmodel <- lm(Income ~ Gender + Role + Age, data = data2)
summary(Linearmodel)
```
